---
title: "Bremer County"
output:
  html_notebook: default
  pdf_document: default
---

Welcome to the Bremer County COVID-19 Data Page.

## Summary
```{r,echo=FALSE}
require(knitr)
colMax <- function(data) sapply(data, max, na.rm = TRUE)

Bremerlastrow <- BremerData[nrow(BremerData),]
Black.Hawklastrow <- Black.HawkData[nrow(Black.HawkData),]
Butlerlastrow <- ButlerData[nrow(ButlerData),]
# AC = as.integer(lastrow["Active"])
# H = as.integer(lastrow["Hospitalized"])
# NP = as.integer(lastrow["New.Pos"])
# PR = as.numeric(round(lastrow["Frac.Pos"]*100, digits = 2))
# ND = as.integer(lastrow["New.Deaths"])
# T = as.integer(lastrow["New.Test"])

temp = data.frame(colMax(BremerData) == Bremerlastrow)
BHtemp = data.frame(colMax(Black.HawkData) == Black.Hawklastrow)
Butemp = data.frame(colMax(ButlerData) == Butlerlastrow)
todaytemp = data.frame(colMax(clean) == today)

###########Risk Calculation#####################
groupsize = 25
Risk = data.frame(Bremer.Prob = Bremerlastrow$Active/Bremer_Population , Black.Hawk.Prob = Black.Hawklastrow$Active/BlackHawk_Population, Butler.Prob = Butlerlastrow$Active/Butler_Population)
Risk = Risk %>% mutate(
  Bremer.Risk = 1-(1-Bremer.Prob*5)^groupsize,
  Black.Hawk.Risk = 1-(1-Black.Hawk.Prob*5)^groupsize,
  Butler.Risk = 1-(1-Butler.Prob*5)^groupsize
  )
```
`r day`: `r as.integer(Bremerlastrow$Active)` active cases `r if(temp$Active) {" (record)"}` in Bremer County, IA today with `r as.integer(Bremerlastrow["Hospitalized"])` hospitalized. If you are in a group of 25 people in Bremer County there is a `r round(Risk$Bremer.Risk*100, digits = 2)`% chance someone in the group is positive.
In the last 24 hours: `r as.integer(Bremerlastrow$New.Pos)` new positive cases`r if(temp$New.Pos) {" (record)"}`, `r as.integer(Bremerlastrow["New.Deaths"])` new death(s) and `r as.integer(today$Bremer.Rep)` tests.

In the last week: `r as.integer(Bremerlastrow["Day7.Pos"])` positives, `r as.integer(Bremerlastrow["Day7.Rec"])` recovered and `r as.integer(Bremerlastrow["Day7.Death"])` death(s). 7 day positivity rate is `r as.integer(today$Bremer.7Day*100)`%.

In the last 14 days: 
 `r as.integer(Bremerlastrow["Day14.Pos"])` positives, `r as.integer(Bremerlastrow["Day14.Rec"])` recovered and `r as.integer(Bremerlastrow["Day14.Death"])` death(s).
 
In our regional medical center for North East Iowa, there are `r as.integer(today["RMCC6.Hospitalized"])` in the hospital`r if(todaytemp$RMCC6.Hospitalized) {" (record)"}` and `r as.integer(today["RMCC6.ICU"])` in the ICU`r if(todaytemp$RMCC6.ICU) {" (record)"}`.
 
 ***
 `r as.integer(Black.Hawklastrow["Active"])` active cases`r if(BHtemp$Active) {" (record)"}` in Black Hawk County, IA today with `r as.integer(Black.Hawklastrow["Hospitalized"])` hospitalized. If you are in a group of 25 people in Black Hawk County there is a `r round(Risk$Black.Hawk.Risk*100, digits = 2)`% chance someone in the group is positive.
In the last 24 hours: `r as.integer(Black.Hawklastrow["New.Pos"])` new positive cases`r if(BHtemp$New.Pos) {" (record)"}`, `r as.integer(Black.Hawklastrow["New.Deaths"])` new death(s) and `r as.integer(Black.Hawklastrow["New.Test"])` new people tested.

In the last week: `r as.integer(Black.Hawklastrow["Day7.Pos"])` positives, `r as.integer(Black.Hawklastrow["Day7.Rec"])` recovered and `r as.integer(Black.Hawklastrow["Day7.Death"])` death(s). 7 day positivity rate is `r as.integer(today$Black.Hawk.7Day*100)`%.

In the last 14 days: 
 `r as.integer(Black.Hawklastrow["Day14.Pos"])` positives, `r as.integer(Black.Hawklastrow["Day14.Rec"])` recovered and `r as.integer(Black.Hawklastrow["Day14.Death"])` death(s).
 
 ***
 `r as.integer(Butlerlastrow["Active"])` active cases`r if(Butemp$Active) {" (record)"}` in Butler County, IA today with `r as.integer(Butlerlastrow["Hospitalized"])` hospitalized`r if(Butemp$Hospitalized) {" (record)"}`. If you are in a group of 25 people in Butler County there is a `r round(Risk$Butler.Risk*100, digits = 2)`% chance someone in the group is positive.
In the last 24 hours: `r as.integer(Butlerlastrow["New.Pos"])` new positive cases, `r as.integer(Butlerlastrow["New.Deaths"])` new death(s) and `r as.integer(Butlerlastrow["New.Test"])` new people tested.

In the last week: `r as.integer(Butlerlastrow["Day7.Pos"])` positives, `r as.integer(Butlerlastrow["Day7.Rec"])` recovered and `r as.integer(Butlerlastrow["Day7.Death"])` death(s). 7 day positivity rate is `r as.integer(today$Butler.7Day*100)`%.

In the last 14 days: 
 `r as.integer(Butlerlastrow["Day14.Pos"])` positives, `r as.integer(Butlerlastrow["Day14.Rec"])` recovered and `r as.integer(Butlerlastrow["Day14.Death"])` death(s).


```{r, echo = FALSE, error=FALSE}
require(reshape2)
mydate <- Sys.Date()

Bremer_Population <- county[county$COUNTY=="Bremer",][,3]
Butler_Population <- county[county$STATE=="Iowa"&county$COUNTY=="Butler",][,3]
BlackHawk_Population <- county[
    county$STATE=="Iowa"&county$COUNTY=="Black Hawk",
    ][,3]

temp = clean[,c("date","Bremer.SS", "Butler.SS", "BlackHawk.SS")]
temp = temp %>% mutate(Bremer = Bremer.SS/Bremer_Population*100000,
                       Butler = Butler.SS/Butler_Population* 100000,
                       BlackHawk = BlackHawk.SS/BlackHawk_Population * 100000)
temp = temp[-(1:71),]
mtemp = melt(temp, id.vars = "date", 
             measure.vars = c('Bremer', 'Butler', 'BlackHawk'))
printplot <- qplot(date, value, color = variable, data = mtemp, geom = c("point", "smooth")) +
    ylab("Cases per 100,000") +
    labs(color = "County", title = paste("Active Cases Per 100,000", mydate))

print(printplot)
png("BBBCounties.png")
printplot
dev.off()
```


Here is the latest information availible as of `r  Sys.Date()` :
```{r data0, echo = FALSE}
# colnames(BremerData) = CountyDataLongNames
lastrow <- BremerData[nrow(BremerData),]
kable(t(lastrow[-1]), digits = 2)
```

```{r data1, echo = FALSE}
kable(BremerData[,1:7])
```


