---
title: "Bremer County"
output:
  html_notebook: default
  pdf_document: default
---

Welcome to the Bremer County COVID-19 Data Page.

## Summary
```{r,echo=FALSE}
require(knitr)
lastrow <- BremerData[nrow(BremerData),]
AC = as.integer(lastrow["Active"])
H = as.integer(lastrow["Hospitalized"])
NP = as.integer(lastrow["New.Pos"])
PR = as.numeric(round(lastrow["Frac.Pos"]*100, digits = 2))
ND = as.integer(lastrow["New.Deaths"])
T = as.integer(lastrow["New.Test"])
```
`r AC` active cases in Bremer County, IA today with `r H` hospitalized and `r ND` new deaths. In the last 24 hours we have had `r NP` new positive tests out of `r T` which gives us a positivity rate of `r  PR`%.

More generally, in the last week, there have been `r as.integer(lastrow["Day7.Pos"])` new cases in the last week with `r as.integer(lastrow["Day7.Pos"])` recoveries. Our 7-day positivity rate is `r as.numeric(round(lastrow["Day7.Perc"]*100, digits = 2))`%. Our 14-day positivity rate is `r as.numeric(round(lastrow["Day14.Perc"]*100, digits = 2))`%.

```{r, echo = FALSE, error=FALSE}
require(reshape2)
mydate <- Sys.Date()

Bremer_Population <- county[county$COUNTY=="Bremer",][,3]
Butler_Population <- county[county$STATE=="Iowa"&county$COUNTY=="Butler",][,3]
BlackHawk_Population <- county[
    county$STATE=="Iowa"&county$COUNTY=="Black Hawk",
    ][,3]

temp = clean[,c("date","Bremer.SS", "Butler.SS", "BlackHawk.SS")]
temp = temp %>% mutate(Bremer = Bremer.SS/Bremer_Population*100000,
                       Butler = Butler.SS/Butler_Population* 100000,
                       BlackHawk = BlackHawk.SS/BlackHawk_Population * 100000)
temp = temp[-(1:71),]
mtemp = melt(temp, id.vars = "date", 
             measure.vars = c('Bremer', 'Butler', 'BlackHawk'))
printplot <- qplot(date, value, color = variable, data = mtemp, geom = c("point", "smooth")) +
    ylab("Cases per 100,000") +
    labs(color = "County", title = paste("Active Cases Per 100,000", mydate))

print(printplot)
png("BBBCounties.png")
printplot
dev.off()
```


Here is the latest information availible as of `r  Sys.Date()` :
```{r data0, echo = FALSE}
# colnames(BremerData) = CountyDataLongNames
lastrow <- BremerData[nrow(BremerData),]
kable(t(lastrow[-1]), digits = 2)
```

```{r data1, echo = FALSE}
kable(BremerData[,1:7])
```


