---
title: "Bremer County"
output:
  html_notebook: default
---

Welcome to the Bremer County COVID-19 Data Page.

## Summary
```{r,echo=FALSE, message=FALSE}
require(knitr)
day <- Sys.Date()
colMax <- function(data) sapply(data, max, na.rm = TRUE)
Bremer_Population <- county[county$COUNTY=="Bremer",][,3]
Butler_Population <- county[county$STATE=="Iowa"&county$COUNTY=="Butler",][,3]
BlackHawk_Population <- county[county$STATE=="Iowa"&county$COUNTY=="Black Hawk",][,3]
Linn_Population <- county[county$STATE=="Iowa"&county$COUNTY=="Linn",][,3]
Johnson_Population <- county[county$STATE=="Iowa"&county$COUNTY=="Johnson",][,3]

Bremerlastrow <- BremerData[nrow(BremerData),]
Black.Hawklastrow <- Black.HawkData[nrow(Black.HawkData),]
Butlerlastrow <- ButlerData[nrow(ButlerData),]
# AC = as.integer(lastrow["Active"])
# H = as.integer(lastrow["Hospitalized"])
# NP = as.integer(lastrow["New.Pos"])
# PR = as.numeric(round(lastrow["Frac.Pos"]*100, digits = 2))
# ND = as.integer(lastrow["New.Deaths"])
# T = as.integer(lastrow["New.Test"])

Brtemp = data.frame(colMax(BremerData) == Bremerlastrow)
BHtemp = data.frame(colMax(Black.HawkData) == Black.Hawklastrow)
Butemp = data.frame(colMax(ButlerData) == Butlerlastrow)
todaytemp = data.frame(colMax(covid19[,-1]) == today[,-1])

################### Find Data Closest to date #########
closestdate <- function(DT, D){
    which(abs(DT$date-(date-D)) == 
              min(abs(DT$date-date+D)))
}

Bremeroneyearago = closestdate(BremerData, 365)
Bremertwoyearago = closestdate(BremerData, 730)

Butleroneyearago = closestdate(ButlerData, 365)
Butlertwoyearago = closestdate(ButlerData, 730)

BHoneyearago = closestdate(Black.HawkData, 365)
BHtwoyearago = closestdate(Black.HawkData, 730)

###########Risk Calculation#####################
groupsize = 30
# Risk <- data.frame(Date = date)

Risk <- data.frame(Date = date,Bremer.Prob = Bremerlastrow$Active/Bremer_Population 
, Black.Hawk.Prob = Black.Hawklastrow$Active/BlackHawk_Population, 
Butler.Prob = Butlerlastrow$Active/Butler_Population,
WSR.Prob = (Butlerlastrow$Active+ Bremerlastrow$Active)/(Butler_Population+ Bremer_Population), 
Linn.Prob = LinnData[nrow(LinnData),]$Active/Linn_Population,
Johnson.Prob = JohnsonData[nrow(JohnsonData),]$Active/Johnson_Population)
names(Risk) = c("Date", "Bremer.Prob", "Black.Hawk.Prob", "Butler.Prob","WSR.Prob", "Linn.Prob", "Johnson.Prob")
Risk = Risk %>% mutate(
  Bremer.Risk = 1-(1-Bremer.Prob*5)^groupsize,
  Black.Hawk.Risk = 1-(1-Black.Hawk.Prob*5)^groupsize,
  Butler.Risk = 1-(1-Butler.Prob*5)^groupsize,
  WSR.Risk = 1-(1-WSR.Prob*5)^groupsize,
  Johnson.Risk = 1-(1-Johnson.Prob*5)^groupsize,
  Linn.Risk = 1-(1-Linn.Prob*5)^groupsize
  )
```
```{r, echo=FALSE}
#  Two years ago, we had `r subset(BremerData, date == (day-730))$Active` active cases and  `r subset(BremerData, date == (day-730))$New.Pos` new positive patients in Bremer County. 
```

`r day`: `r as.integer(Bremerlastrow$Active)` active cases `r if(Brtemp$Active) {" (record)"}` in Bremer County, IA today. If the current infection trends continue, in 4 weeks we will have `r as.integer(Bremerlastrow["New.Pos"]*4)` active cases.

In the last 7 days: `r New.Positive.Tests[nrow(New.Positive.Tests),"Bremer"]` new positive tests, `r as.integer(Bremerlastrow$New.Pos)`  new positive cases`r if(Brtemp$New.Pos) {" (record)"}`,  and `r as.integer(Bremerlastrow["New.Deaths"])` new death(s).

If you are in a group of `r groupsize` random people in Bremer County there is a `r round(Risk$Bremer.Risk*100, digits = 2)[1,]`% chance someone in the group is positive. One in `r round(Bremer_Population/Bremerlastrow$Deaths, digits = 0)[1,]` people in Bremer County has died of COVID-19.

 One year ago, we had `r BremerData[Bremeroneyearago,]$Active` active cases and  `r BremerData[Bremeroneyearago,]$New.Pos` new positive per day patients in Bremer County. 
 
 Two years ago, we had `r BremerData[Bremertwoyearago,]$Active` active cases and  `r BremerData[Bremertwoyearago,]$New.Pos` new positive per day patients in Bremer County. 

***
 `r as.integer(Butlerlastrow["Active"])` active cases`r if(Butemp$Active) {" (record)"}` in Butler County, IA today. If the current infection trends continue, in 4 weeks we will have `r as.integer(Butlerlastrow["New.Pos"]*4)` active cases. 
 
In the last 7 days: `r New.Positive.Tests[nrow(New.Positive.Tests),"Butler"]` new positive tests, `r as.integer(Butlerlastrow["New.Pos"])` new positive cases,  and `r as.integer(Butlerlastrow["New.Deaths"])` new death(s).

If you are in a group of `r groupsize` people in Butler County there is a `r round(Risk$Butler.Risk*100, digits = 2)[1,]`% chance someone in the group is positive. One in `r round(Butler_Population/Butlerlastrow$Deaths, digits = 0)[1,]` people in Butler County has died of COVID-19.

 One year ago, we had `r ButlerData[Butleroneyearago,]$Active` active cases and  `r ButlerData[Butleroneyearago,]$New.Pos` new positive per day patients in Butler County. 
 
 Two years ago, we had `r ButlerData[Butlertwoyearago,]$Active` active cases and  `r ButlerData[Butlertwoyearago,]$New.Pos` new positive per day patients in Butler County. 
 
 ***
 `r as.integer(Black.Hawklastrow["Active"])` active cases`r if(BHtemp$Active) {" (record)"}` in Black Hawk County, IA today. If the current infection trends continue, in 4 weeks we will have `r as.integer(Black.Hawklastrow["New.Pos"])*4` active cases. If you are in a group of `r groupsize` people in Black Hawk County there is a `r round(Risk$Black.Hawk.Risk*100, digits = 2)[1,]`% chance someone in the group is positive. One in `r round(BlackHawk_Population/Black.Hawklastrow$Deaths, digits = 0)[1,]` people in Black Hawk County has died of COVID-19.
 
In the last 7 days: `r New.Positive.Tests[nrow(New.Positive.Tests),"Black.Hawk"]` new positive tests, `r as.integer(Black.Hawklastrow["New.Pos"])` new positive cases`r if(BHtemp$New.Pos) {" (record)"}`, and `r as.integer(Black.Hawklastrow["New.Deaths"])` new death(s).

 One year ago, we had `r Black.HawkData[BHoneyearago,]$Active` active cases and  `r Black.HawkData[BHoneyearago,]$New.Pos` new positive per day patients in Black Hawk County. 
 
 Two years ago, we had `r Black.HawkData[BHtwoyearago,]$Active` active cases and  `r Black.HawkData[BHtwoyearago,]$New.Pos` new positive per day patients in Black Hawk County. 

 
```{r, echo = False, message=FALSE}
require(reshape2)
mydate <- Sys.Date()
graphlength = 90

Bremer_Population <- county[county$COUNTY=="Bremer",][,3]
Butler_Population <- county[county$STATE=="Iowa"&county$COUNTY=="Butler",][,3]
BlackHawk_Population <- county[
    county$STATE=="Iowa"&county$COUNTY=="Black Hawk",
    ][,3]

year_reference = subset(BremerData, date == (day-365))$Active/Bremer_Population*10^5

temp = covid19[,c("date","Bremer.SS", "Butler.SS", "BlackHawk.SS")]
temp = temp %>% mutate(Bremer = Bremer.SS/Bremer_Population$Population*100000,
                       Butler = Butler.SS/Butler_Population$Population* 100000,
                       BlackHawk = BlackHawk.SS/BlackHawk_Population$Population * 100000)
# temp = tail(temp, n=graphlength)
temp = temp[temp$date >= (mydate-graphlength),]
mtemp = melt(temp, id.vars = "date", 
             measure.vars = c('Bremer', 'Butler', 'BlackHawk'))
printplot <- qplot(as.Date(date), value, color = variable, data = mtemp, geom = "point") +
  geom_smooth(span = 0.5, se = TRUE) +
    ylab("Cases per 100,000") +
    xlab("Date")+
    labs(color = "County", title = paste("Active Cases Per 100,000", mydate))
  # +geom_hline(yintercept = year_reference, 
  #            color = "red") + 
  # geom_text(
  #   aes(x= temp$date[1]+20, y= year_reference, 
  #       label = day-365, vjust = -0.5), 
  #   colour = "black")

print(printplot)
ggsave(paste(day,"BBBCounties.png"), plot = printplot)
# ggsave("BBBCounties.png")
# 
# png(paste(day,"BBBCounties.png"))
# print(printplot)
# dev.off()
```


Here is the latest information availible as of `r  Sys.Date()` :
```{r data0, echo = FALSE}
# colnames(BremerData) = CountyDataLongNames
lastrow <- BremerData[nrow(BremerData),]
kable(t(lastrow[-1]), digits = 2)
```

```{r data1, echo = FALSE}
kable(BremerData[,1:7])
```

