---
title: "Covid-19 Analysis"
author: "Mariah Birgen"
output: html_notebook
---

```{r initialization, echo=FALSE}
suppressPackageStartupMessages( 
  require(dplyr))
suppressPackageStartupMessages( require(lubridate))
suppressPackageStartupMessages( require(ggplot2))
suppressPackageStartupMessages( require(zoo))
suppressPackageStartupMessages( require(gsheet))
suppressPackageStartupMessages( require(plyr))
suppressPackageStartupMessages( require(pdftools))
suppressPackageStartupMessages( require(tidyverse))
suppressPackageStartupMessages( require(stringr))
# suppressPackageStartupMessages( require(readxl))
```

Load data
```{r, echo=FALSE}
mydate1 <- Sys.Date()
hosp_by_county <- pdf_text("hosp_by_county.pdf") %>%
    readr::read_lines() %>% str_squish()
# hosp_by_county <-    str_squish(hosp_by_county)
temp <- hosp_by_county
hosp_by_county <- temp[5:37]
temp1 = strsplit(temp[40]," ") %>% ldply()
date = as.character(mdy(temp1[3]))

hosp_by_county <-hosp_by_county %>%
    strsplit("(?= [A-Za-z])(?<=[0-9])|(?= [0-9])(?<=[A-Za-z])", perl=TRUE)

hosp_by_county_df <- plyr::ldply(hosp_by_county)
temp1 <- hosp_by_county_df[,1:2]
temp2 <- hosp_by_county_df[,3:4]
temp3 <- hosp_by_county_df[,5:6]
colnames(temp2) <- names(temp1)
colnames(temp3) <- names(temp1)
hosp_by_county <- rbind(temp1, temp2, temp3)
colnames(hosp_by_county) <- c("County", "Hospitalized")
hosp_by_county <-hosp_by_county %>%
    mutate(Hospitalized = as.numeric(Hospitalized))
temp1 = sum(hosp_by_county[,2])
hosp_by_county <- rbind(hosp_by_county, c("Totals", temp1))
rm(temp1, temp2, temp3, hosp_by_county_df)

rownames(hosp_by_county) <- hosp_by_county[,1]

hospital <- as.data.frame(read.csv("CountyHospitilizations.csv"))
hospital <- hospital[,-1]
#hospital[,1] = mdy(hospital[,1])
h_county <-  t(hosp_by_county[,2])

colnames(h_county) <-  hosp_by_county[,1]
h_county <- cbind(date = c(date, h_county))
rownames(h_county) <- colnames(hospital)
h_county <- t(h_county)
if(mydate1== date){
    hospital <- rbind(hospital, h_county)
} else {
  temp <-hospital[,1]== date
  temp1 <- (1:600)[temp][1]
  hospital[temp1,] <- h_county
  if(as.character(hospital[nrow(hospital),1]) != mydate1){
  hospital <- rbind.fill(hospital, data.frame(Date = Sys.Date()))
  }
}


write.csv(hospital, "CountyHospitilizations.csv")
hospital <- hospital[!is.na(hospital$Adair),]
```

```{r}
Summary <- pdf_text("Iowa COVID-19 Information - access.pdf") %>%
    readr::read_lines() %>% str_squish()
# temp1 = strsplit(Summary[1]," ") %>% ldply()
temp1 = Summary[6] %>% ldply()
date = as.character(mdy(temp1[1]))

Summary <- str_remove_all(Summary, " Translate") %>% 
    str_remove_all("Translate")
CountyDataOld <- Summary[44:149] %>%
    strsplit("(?= [A-Za-z])(?<=[0-9])|(?= [0-9])(?<=[A-Za-z])", perl=TRUE)
CountyDataOld <- CountyDataOld[-c(10:12,72:74)]
CountyDataOld <- plyr::ldply(CountyDataOld)
temp <- strsplit(CountyDataOld[,2], " ", fixed = TRUE)
temp <- plyr::ldply(temp)
temp[,1] = CountyDataOld[,1]
colnames(temp) = c("County", "Individuals.Tested","Individuals.Positive", 
                   "Total.Recovered", "Total.Deaths")
CountyDataOld <- temp
rm(temp)
CountyDataOld[-1] = sapply(CountyDataOld[-1], as.integer)
CountyDataOld <- as.data.frame(CountyDataOld) %>%
    arrange(County)%>%
    mutate(Active = Individuals.Positive - Total.Recovered - Total.Deaths,
           positivity = Individuals.Positive/Individuals.Tested,
           date = as.Date(mydate),
    )

temp <-c(County = "Totals",
          colSums(CountyDataOld[,2:6]),
          positivity = NA,
          date = as.character(date))
temp["positivity"] = as.integer(temp["Individuals.Positive"])/as.integer(temp["Individuals.Tested"])
CountyDataOld[101,] = temp
CountyDataOld[,2:6] = sapply(CountyDataOld[,2:6], as.integer)
CountyDataOld[,7] = sapply(CountyDataOld[,7], as.numeric)
```

```{r, echo=FALSE}
mydate <- Sys.Date()

filename <- "censusCounties.csv"
county <- read.csv(filename,
                       stringsAsFactors = FALSE)

filename <- "Summary.csv"
CountyData <- read.csv(filename,
                       stringsAsFactors = FALSE)
CountyData <- as.data.frame(CountyData) %>%
    arrange(County) %>%
  mutate(Individuals.Tested = CountyDataOld$Individuals.Tested[-101], Individuals.Positive= CountyDataOld$Individuals.Positive[-101]) %>%
    mutate(Active = Individuals.Positive - Total.Recovered - Total.Deaths,
           positivity = Total.Positive.Tests/Total.Tests,
           date = as.Date(mydate),
           )


if(CountyData[1,1]== ""){
  CountyData = CountyData[-1,]
}
rownames(CountyData) <- CountyData[,1]
CountyData[,2:10] = sapply(
  CountyData[,2:10], as.integer)
CountyData[,11] = sapply(CountyData[,11], as.numeric)
temp <-c( County = "Totals",
          colSums(CountyData[,2:10]),
          positivity = NA,
          date = as.character(mydate))
temp["positivity"] = as.integer(temp["Individuals.Positive"])/as.integer(temp["Individuals.Tested"])
CountyData["Total",] = temp
CountyData[,2:10] = sapply(
  CountyData[,2:10], as.integer)
CountyData[,11] = sapply(CountyData[,11], as.numeric)
# CountyData <- rbind(CountyData, Total = temp)
CountyData <-cbind(date = as.Date(mydate), CountyData[,-11])
write.csv(CountyData, file = paste(
    'CountyData/',mydate,'CountyData.csv'))

d_county <- t(CountyData[,-1])
county_names <- d_county[1,]
colnames(d_county) <- d_county[1,]
d_county = d_county[-1,]
# colnames(d_county)[1] <- "date"
d_county[-10,] = sapply(d_county[-10,], as.integer)
# d_county[9,] = sapply(d_county[9,], as.numeric)
# d_county <- as.data.frame(d_county)%>%
#     mutate(Total = Reduce(`+`, .))
d_county <- cbind(date=as.character(mydate),d_county)

Tested <- read.csv( "CountyData/CountyTests.csv", stringsAsFactors=FALSE)
Tested <- Tested[,-1]
Positive <- read.csv( "CountyData/CountyPositive.csv", stringsAsFactors=FALSE)
Positive <- Positive[,-1]
Recovered <- read.csv( "CountyData/CountyRecovered.csv", stringsAsFactors=FALSE)
Recovered <- Recovered[,-1]
Deaths <- read.csv( "CountyData/CountyDeaths.csv", stringsAsFactors=FALSE)
Deaths <- Deaths[,-1]
Active <- read.csv( "CountyData/CountyActive.csv", stringsAsFactors=FALSE)
Active <- Active[,-1]
templist = c("Tested", "Positive", "Recovered", "Deaths", "Active" )


if( as.character(Tested[nrow(Tested), 1]) != mydate){
Tested <- rbind(Tested, d_county["Individuals.Tested",])
Positive <- rbind(Positive, d_county["Individuals.Positive",])
Recovered <- rbind(Recovered, d_county["Total.Recovered",])
Deaths <- rbind(Deaths, d_county["Total.Deaths",])
Active <- rbind(Active, d_county["Active",])
}

Tested[,-1] = sapply(Tested[,-1], as.integer)
Positive[,-1] = sapply(Positive[,-1], as.integer)
Recovered[,-1] = sapply(Recovered[,-1], as.integer)
Deaths[,-1] = sapply(Deaths[,-1], as.integer)
Active[,-1] = sapply(Active[,-1], as.integer)

trows <- nrow(Tested)
newTest <- Tested[trows,-1]-Tested[trows-1,-1]
newPos <- Positive[trows,-1]-Positive[trows-1,-1]
newRec <- Recovered[trows,-1]-Recovered[trows-1,-1]
newDeath <- Deaths[trows,-1]-Deaths[trows-1,-1]
PerPos7 <- (Positive[trows,-1]-
              Positive[trows-7,-1])/
  (Tested[trows,-1]-
     Tested[trows-7,-1])

names_list <- rownames(CountyData)
names_list <- names_list[-c(75,101)]
# names_list <- names_list[-100]
pop <- c(1:99)

per100 <- Tested[,1]
for(i in names_list){
  pop[i] <- county[
    county$STATE=="Iowa"&county$COUNTY==i,][,3]
  tempname <- gsub(" ", "." , i)
  tempname = gsub("'",".", tempname)
  per100 <- cbind(per100,Active[tempname]/pop[i]*100000)
}

county_names <- colnames(Tested)
county_names <- county_names[-76]
county_names <- gsub(" ", ".", county_names)
for(i in 2:(length(county_names))){
  i = county_names[i]
    test <- Tested[,i]
    temp = length(test)
    pos <- as.numeric(Positive[,i])
    rec <- as.numeric(Recovered[,i])
    hosp <- as.integer(hospital[,i])
    hosp = c(rep(NA, length(test)-length(hosp)),hosp)
    death <- as.numeric(Deaths[,i])
    dates <- as.Date(Tested[,1]    )
    temp <- data.frame(
        date = dates,
        Tested = test,
        Positive = pos,
        Recovered = rec,
        Hospitalized = hosp,
        Deaths = death)
    
    temp <- temp %>% 
        mutate(
            Active = Positive - Recovered - Deaths,
            New.Pos = Positive - lag(Positive),
            New.Test = Tested - lag(Tested),
            New.Rec = Recovered - lag(Recovered),
            New.Deaths = Deaths - lag(Deaths),
            Frac.Pos = New.Pos/New.Test,
            Day7.Test = Tested - lag(Tested, n=7),
            Day7.Pos = Positive - lag(Positive, n=7),
            Day7.Rec = Recovered - lag(Recovered, n=7),
            Day7.Death = Deaths - lag(Deaths, n=7),
            Day7.Active = Active - lag(Active, n=7),
            Day7.Perc = Day7.Pos/Day7.Test,
            Day14.Test = Tested - lag(Tested, n=14),
            Day14.Pos = Positive - lag(Positive, n=14),
            Day14.Rec = Recovered - lag(Recovered, n=14),
            Day14.Death = Deaths - lag(Deaths, n=14),
            Day14.Active = Active - lag(Active, n=14),
            Day14.Perc = Day14.Pos/Day14.Test,
            Not.Sick = (lag(Positive, n= 10) -
                          lag(Positive, n=28))*0.7,
            True.Active = Positive - Recovered -
              Deaths - round(Not.Sick, digits = 0)
        )
    
    assign(paste(i,"Data", sep=""), temp)
    # write.csv(temp, paste("CountyData/",county_names[i],
                          # "Data.csv", sep=""))
}

write.csv(Tested, "CountyData/CountyTests.csv")
write.csv(Positive, "CountyData/CountyPositive.csv")
write.csv(Recovered, "CountyData/CountyRecovered.csv")
write.csv(Deaths, "CountyData/CountyDeaths.csv")
write.csv(Active, "CountyData/CountyActive.csv")

```


```{r load, echo = FALSE}
colMax <- function(data) sapply(data, max, na.rm = TRUE)

covid19 <- read.csv("covid19.csv", stringsAsFactors = FALSE)
# covid19 <- read_excel("covid19.xlsx", sheet = 1)

county <- read.csv("censusCounties.csv", stringsAsFactors = FALSE)
# covid19 = covid19[-158,]
names(covid19)[names(covid19) == "ï..date"] <- 'date'
# covid19$date <- mdy(covid19$date)
covid19$date <- ymd(covid19$date)
covid19$deaths = sapply(covid19$deaths, as.integer)

covid19 <-covid19 %>% mutate(
  negative = Total.Tested - positive,
  New.Positive = positive - lag(positive),
  New.Negative = negative - lag(negative),
  new.hospital = hospitalized - lag(hospitalized),
  Total.Daily.Tests = Total.Tested - lag(Total.Tested),
  New.Deaths = deaths - lag(deaths),
  Still.Sick = positive - Recovered) %>% mutate(
    New.Percent.Positive = New.Positive/Total.Daily.Tests,
    Week.Positivity.Rate = (positive - lag(positive, n=7))/(Total.Tested - lag(Total.Tested, n=7)),
    Percent.Hospitalized = hospitalized/Still.Sick
  )
#rownames(covid19) <- covid19[,"date"]

#################################################
##  Fill in county Data in covid19
#################################################
numrows = nrow(BremerData)
numrows1 = nrow(Black.HawkData)
tempBr =   BremerData[numrows,
                      c("Positive", "Recovered","Deaths", "Active" )]
tempBl =   Black.HawkData[numrows1,
                      c("Positive", "Recovered","Deaths", "Active" )]
tempBu =   ButlerData[numrows1,
                      c("Positive", "Recovered","Deaths", "Active" )]
names(tempBr) = c("Bremer.Positive", "Bremer.Recovered",
                  "Bremer.Death", "Bremer.SS") 
names(tempBu) = c("Butler.P", "Butler.R",
                  "Butler.D", "Butler.SS") 
names(tempBl) = c("BlackHawk.P", "BlackHawk.R",
                  "BlackHawk.D", "BlackHawk.SS") 
temp = cbind(tempBr, tempBu, tempBl)
for (i in names(temp)){
      covid19[nrow(covid19),i] = temp[i]
    }

write.csv(covid19, "covid19.csv",row.names = FALSE)
#################################################
##  Create Clean Dataset
#################################################

clearn <- !is.na(covid19$positive)
clean <- covid19[clearn,]
cdeaths <- covid19[!is.na(covid19$deaths),]
clean <- clean %>% mutate(
  negative = Total.Tested - positive,
  Percent.Pos = positive/Total.Tested*100,
  #np_7day = rollmean(New.Positive, k=7, align = "right", fill = NA),
  np_7day = (positive - lag(positive, n=7))/7,
  New.Percent.Positive = New.Positive/Total.Daily.Tests,
  Percent.Hospitalized = hospitalized/Still.Sick*100
)
clean <- clean %>% mutate(
#  np_7day = rollmean(New.Positive, k=7, align = "right", fill = NA),
  #Week.Positivity = rollapply(New.Positive, 7, FUN=sum, fill = NA),
  #Week.Tests = rollapply(Total.Daily.Tests, 7, FUN=sum, fill = NA),
  Week.Positivity = positive - lag(positive, n=7),
  Week.Tests = Total.Tested - lag(Total.Tested, n=7),
  Week.Positivity.Rate = Week.Positivity/Week.Tests,
  Week.Deaths = deaths - lag(deaths, n=7),
  Day.Total.Tests = Total.Tests - lag(Total.Tests,n=1),
  Day.Positive.Tests = Positive.Tests - lag(Positive.Tests, n=1),
  Day.Bremer.Total.Tests = Bremer.Total.Tests - lag(Bremer.Total.Tests,n=1),
  Day.Bremer.Positive.Tests = Bremer.Positive.Tests - lag(Bremer.Positive.Tests, n=1),
  )
clean <- clean %>% mutate(wktot = Week.Positivity)
# clean$wktot <- sapply(clean$date,function(x)sum(clean[difftime(clean$date,x,,"days") %in% 0:-6,]$New.Positive))
lastrow <- nrow(clean)
clean$deaths = sapply(clean$deaths, as.integer)
clean <- clean %>% mutate(
  nd_7day = deaths - lag(deaths, n=7), #rollmean(deaths, k=7, fill = NA,),
  pp_7day = (positive - lag(positive, n=7))/ (Total.Tested - lag(Total.Tested, n=7))#
    #rollmean(New.Percent.Positive, k=7, fill = NA)
    )
```
```{r, echo = FALSE}
today <- clean[nrow(clean),]
day <- today$date
new_pos <- today$New.Positive
percent_pos <- round(clean$New.Percent.Positive[nrow(clean)]*100, digits = 1)
# percent_pos <- today$New.Percent.Positive
percent_pos7 <- round(clean$Week.Positivity.Rate[nrow(clean)-4]*100, digits = 1)

percent_hospital <- round(clean$Percent.Hospitalized[nrow(clean)], digits = 1)
# percent_hospital <- clean$Percent.Hospitalized[nrow(clean)]
ICU <- today$ICU
percent_ICU <- round(today$ICU/today$hospitalized*100, digits = 1)
sick <- format(today$Still.Sick, scientific = FALSE, big.mark=",")
IowaPop <- 3155070

clean = clean %>% mutate(
  Individuals.Reported = Individuals.Tested - lag(Individuals.Tested),
  Individuals.Positive = Individuals.Postive - lag(Individuals.Postive),
  Bremer.Rep = Bremer.Individuals.Tested - lag(Bremer.Individuals.Postive))

# Adding more data to Bremer
hosp <- as.integer(hospital[,"Bremer"])
temp <- clean %>% select("date","Bremer.Positive", 
                         "Bremer.Recovered", "Bremer.Death")
names(temp) = c("date","Positive", "Recovered", "Deaths")
temp <- subset(temp, 
               !is.na(Positive) & date < "2020-07-31")%>% 
  mutate(Tested = NA, Hospitalized = NA)
temp2 <- BremerData %>% select("date","Hospitalized", "Tested","Positive", "Recovered", "Deaths")
temp = rbind(temp, temp2)
rm(temp2)
temp$Tested = as.integer(temp$Tested)
temp <- temp  %>% 
  mutate(
    Active = Positive - Recovered - Deaths,
    New.Pos = Positive - lag(Positive),
    New.Test = Tested - lag(Tested),
    New.Rec = Recovered - lag(Recovered),
    New.Deaths = Deaths - lag(Deaths),
    Frac.Pos = New.Pos/New.Test,
    Day7.Test = Tested - lag(Tested, n=7),
    Day7.Pos = Positive - lag(Positive, n=7),
    Day7.Rec = Recovered - lag(Recovered, n=7),
    Day7.Death = Deaths - lag(Deaths, n=7),
    Day7.Active = Active - lag(Active, n=7),
    Day7.Perc = Day7.Pos/Day7.Test,
    Day14.Test = Tested - lag(Tested, n=14),
    Day14.Pos = Positive - lag(Positive, n=14),
    Day14.Rec = Recovered - lag(Recovered, n=14),
    Day14.Death = Deaths - lag(Deaths, n=14),
    Day14.Active = Active - lag(Active, n=14),
    Day14.Perc = Day14.Pos/Day14.Test,
            Not.Sick = (lag(Positive, n= 10) -
                          lag(Positive, n=28))*0.7,
            True.Active = Positive - Recovered -
              Deaths - round(Not.Sick, digits = 0)
  )

BremerData = temp
rm(temp)
write.csv(BremerData, "CountyData/BremerData.csv")

maxtoday = data.frame(colMax(clean[,-1]) == today[,-1])
```
# Iowa COVID-19 Summary `r day`

Today we had `r new_pos` new positive patients `r if(maxtoday$New.Positive) {" (record)"}` which is `r percent_pos`% (`r percent_pos7`% in the last week) of the newly tested people recorded in the last 24 hours.  `r clean$Week.Deaths[nrow(clean)]` deaths in the last week`r if(maxtoday$Week.Deaths) {" (record)"}`. One in `r round(IowaPop/clean[nrow(clean),]$deaths, digits = 0)` people in Iowa has died of COVID-19. There are `r sick` people actively sick`r if(maxtoday$Still.Sick) {" (record)"}` and `r percent_hospital`% (`r today$hospitalized`) are hospitalized`r if(maxtoday$hospitalized) {" (record)"}`. `r percent_ICU`% (`r ICU`) of those hospitalized are in the ICU`r if(maxtoday$ICU) {" (record)"}`. 

#Print Graphs

 ```{r, echo=FALSE}
pactive <- qplot(date, Still.Sick, data = clean[-(1:23),], geom = c("point", "smooth")) +
ggtitle("Known Active COVID-19 Cases in Iowa")
png(paste(day,"pactive.png"))
pactive
dev.off()
 pactive
 ```

 ## Percent Positive Active Cases
 ```{r,echo=False}

 ppositive <-qplot(date, New.Percent.Positive, data = clean[-(1:7),], geom = c("point", "smooth")) + ylab("Fraction Positive") +
 ggtitle("Fraction of Positive COVID-19 Tests in Iowa")
 png(paste(day,"ppositive.png"))
 ppositive
 dev.off()
 ppositive

 p7positive <-qplot(date, Week.Positivity.Rate, data = clean[-(1:20),], geom = c("point", "smooth")) + ylab("Fraction Positive") +
 ggtitle("7 day average of Fraction of Positive COVID-19 Tests in Iowa")
 png(paste(day,"p7positive.png"))

 p7positive
 dev.off()
 p7positive
 ```


 ## Bremer County
 I started recording daily data on Bremer county on May 29, 2020. There will be problems with the recovered data since I have yet to get tables of data for the county and the definition of recovered changed on June 30. Thus, the following graph omits that data.
 ```{r, echo=FALSE}
 State_Average_Active <- today$Still.Sick/IowaPop*100000
 Bremer_Population <- county[county$COUNTY=="Bremer",][,3]
 Butler_Population <- county[county$STATE=="Iowa"&county$COUNTY=="Butler",][,3]
 BlackHawk_Population <- county[county$STATE=="Iowa"&county$COUNTY=="Black Hawk",][,3]
 pBremer <-qplot(date,
                 Bremer.SS/Bremer_Population*100000,
                 Bremer.SS,
                 data = clean[-(1:71),], geom = c("point", "smooth")) +
   ggtitle(
 "Known Active COVID-19 Cases in Bremer County Per 100,000 People")
 #    "Known Active COVID-19 Cases in Bremer County")
 #+  geom_hline(aes(yintercept=State_Average_Active)
 pButler <-qplot(date,
                 #Butler.SS/Butler_Population*100000,
                 Butler.SS,
                 data = clean[-(1:116),], geom = c("point", "smooth")) +
   ggtitle("Known Active COVID-19 Cases in Butler County")
ggtitle("Known Active COVID-19 Cases in Butler County Per 100,000 People")
#+  geom_hline(aes(yintercept=State_Average_Active)
 png(paste(day,"pButler.png"))
 pButler
 dev.off()
 pButler
 pBlackHawk <-qplot(date,
               BlackHawk.SS,
               BlackHawk.SS/BlackHawk_Population*100000,
               data = clean[-(1:116),], geom = c("point", "smooth")) +
   ggtitle("Known Active COVID-19 Cases in Black Hawk County")
 #ggtitle("Known Active COVID-19 Cases in Black Hawk County Per 100,000 People")
 #+  geom_hline(aes(yintercept=State_Average_Active)
 png(paste(day,"pBlackHawk.png"))
 pBlackHawk
 dev.off()
 pBlackHawk

 png(paste(day,"pBremer.png"))
 pBremer
 dev.off()
 pBremer

 day7pos.Bremer <- round(
   BremerData[nrow(BremerData), ncol(BremerData)]*100, digits=1)
 ```