---
title: "Iowa Hospital Data"
output: html_notebook
---

```{r storedata1, echo=FALSE}
date = Sys.Date()
covid19 <- read.csv("covid19.csv", 
                    stringsAsFactors = FALSE)
names(covid19)[1] <- "date"
require(anytime)
closestdate <- function(DT, D){
    which(abs(anydate(DT$date)-(date-D)) == 
              min(abs(anydate(DT$date)-date+D)))
}

covid19$date = anydate(covid19$date)
oneyearago = closestdate(covid19,365)
twoyearago = closestdate(covid19,730)
oneweekago = closestdate(inpatientSave,7)
tail_oneweekago = nrow(inpatientSave) - oneweekago
tail_onemonthago = nrow(inpatientSave) - closestdate(inpatientSave,28)

# inpatientSave = inpatientSave[-nrow(inpatientSave),]
source("HospitalData.R")
inpatientSave <- inpatientSave[!duplicated(inpatientSave[,-1]),]


tail(inpatientSave[,c("date","inpatient_beds_used_covid")],7)
```

```{r}
covid19$date = anydate(covid19$date)
# qplot(date, hospitalized, 
ggplot(data = tail(covid19, n=80), aes(date, hospitalized)) + geom_point() +geom_smooth(span = 0.3, se = TRUE) 

##############################################
mergedYearonYearHospMeans = covid19 %>%
    drop_na(hospitalized) %>%
    select("date","hospitalized") %>%
    mutate(week = week(date),
           year = as.factor(year(date))) %>%
    group_by(year, week) %>%
    summarise_at(vars(hospitalized), list(WeeklyAve = mean)) 

ggplot(data=mergedYearonYearHospMeans, aes(week, WeeklyAve, color = year)) + geom_point()+geom_smooth(span = 0.5, se = TRUE)  + xlab("Week of Year")+ ylab("Weekly Average Hospitalized")
```

## Iowa COVID-19 Hospitalization Summary `r date`


According to HHS, there are `r inpatientSave[nrow(inpatientSave),"inpatient_beds_used_covid"]`  Iowans hospitalized for COVID-19, `r tail(inpatientSave$total_adult_patients_hospitalized_confirmed_and_suspected_covid, n=1)` adults and `r tail(inpatientSave$total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, n=1)` children. Previous day admission was `r tail(inpatientSave$previous_day_admission_adult_covid_confirmed, n=1) + tail(inpatientSave$previous_day_admission_adult_covid_suspected,1) ` adults and `r tail(inpatientSave$previous_day_admission_pediatric_covid_confirmed, n=1) + tail(inpatientSave$previous_day_admission_pediatric_covid_suspected,1) ` children. Of those, `r tail(inpatientSave$total_ICU, n=1)` (`r tail( round(inpatientSave$total_ICU/inpatientSave$inpatient_beds_used_covid*100,1) , n=1)`% of hospitalized) are in the ICU, `r tail(inpatientSave$staffed_icu_adult_patients_confirmed_covid, n=1)` adults and `r tail(inpatientSave$staffed_icu_pediatric_patients_confirmed_covid, n=1)` children. Hospitals have reported `r sum(tail(inpatientSave$deaths_covid, n=tail_oneweekago))` COVID-19 deaths in the last 7 days and `r sum(tail(inpatientSave$deaths_covid, n=tail_onemonthago))` COVID-19 deaths in the last 4 weeks.