---
title: "Covid-19 Analysis"
author: "Mariah Birgen"
output: html_notebook
---

```{r initialization, results='hide', echo=FALSE}
 
 library(dplyr)
 library(lubridate)
 library(ggplot2)
 library(zoo)
 library(gsheet)
 library(plyr)
 library(pdftools)
 library(tidyverse)
 library(stringr)
 library(anytime)
#  require(readxl)

colMax <- function(data) sapply(data, max, na.rm = TRUE)
```

Load data
```{r loadcounty, results='hide', echo=FALSE}
library(readr)
county = read_csv("censusCounties.csv")
## Select data for only Iowa Counties
county = subset(county, STATE == "Iowa")
```


```{r hospitalization, echo=FALSE}
# mydate <- Sys.Date()
# hosp_by_county <- pdf_text("hosp_by_county.pdf") %>%
#     readr::read_lines() %>% str_squish()
# # hosp_by_county <-    str_squish(hosp_by_county)
# temp <- hosp_by_county
# hosp_by_county <- temp[5:37]
# temp1 = strsplit(temp[41]," ") %>% ldply()
# date = as.character(mdy(temp1[3]))
# 
# hosp_by_county <-hosp_by_county %>%
#     strsplit("(?= [A-Za-z])(?<=[0-9])|(?= [0-9])(?<=[A-Za-z])", perl=TRUE)
# 
# hosp_by_county_df <- plyr::ldply(hosp_by_county)
# temp1 <- hosp_by_county_df[,1:2]
# temp2 <- hosp_by_county_df[,3:4]
# temp3 <- hosp_by_county_df[,5:6]
# colnames(temp2) <- names(temp1)
# colnames(temp3) <- names(temp1)
# hosp_by_county <- rbind(temp1, temp2, temp3)
# colnames(hosp_by_county) <- c("County", "Hospitalized")
# hosp_by_county <-hosp_by_county %>%
#     mutate(Hospitalized = as.numeric(Hospitalized))
# temp1 = sum(hosp_by_county[,2])
# hosp_by_county <- rbind(hosp_by_county, c("Totals", temp1))
# rm(temp1, temp2, temp3, hosp_by_county_df)
# 
# rownames(hosp_by_county) <- hosp_by_county[,1]
# 
# hospital <- as.data.frame(read.csv("CountyHospitilizations.csv"))
# hospital <- hospital[,-1]
# #hospital[,1] = mdy(hospital[,1])
# h_county <-  t(hosp_by_county[,2])
# 
# colnames(h_county) <-  hosp_by_county[,1]
# h_county <- cbind(date = c(date, h_county))
# rownames(h_county) <- colnames(hospital)
# h_county <- t(h_county)
# # if(mydate1== date){
#     hospital <- rbind(hospital, h_county)
# # } else {
# #   temp <-hospital[,1]== date
# #   temp1 <- (1:600)[temp][1]
# #   hospital[temp1,] <- h_county
# #   if(as.character(hospital[nrow(hospital),1]) != mydate1){
#   # hospital <- rbind.fill(hospital, data.frame(Date = date))
# #   }
# # }
# hospital <- unique(hospital)
# 
# write.csv(hospital, "CountyHospitilizations.csv")
# hospital <- hospital[!is.na(hospital$Adair),]
```

```{r accessdata}
## Change Date on pdf file
Summary <- pdf_text(
    "access _ Iowa COVID-19 Information25.pdf") %>%
    readr::read_lines() %>% str_squish()
Summary = Summary[!Summary == ""]
temp1 = Summary[6] %>% ldply() 
date = as.character(mdy(temp1[1]))
# Summary[79] = paste(Summary[79], Summary[80])
Summary <- str_remove_all(Summary, " Translate") %>% 
    str_remove_all("Translate")
# Summary[79] = paste(Summary[79], Summary[80])
CountyDataOld <- Summary[46:150]
CountyDataOld <- CountyDataOld %>% 
strsplit("(?= [A-Za-z])(?<=[0-9])|(?= [0-9])(?<=[A-Za-z])", perl=TRUE)


# CountyDataOld <- Summary[46:151] %>%
#     strsplit("(?= [A-Za-z])(?<=[0-9])|(?= [0-9])(?<=[A-Za-z])", perl=TRUE)
# CountyDataOld <- CountyDataOld[-c(10:11,71:72)]
# temp1 = CountyDataOld[44:45]
# temp2 = plyr::ldply(temp1[1])
# temp2[,2]=paste0(temp2[,2],plyr::ldply(temp1[2]) )
# CountyDataOld <- CountyDataOld[-c(35:37,89,91:92)]
CountyDataOld <- CountyDataOld[-c(35:36,88, 90:91)]
CountyDataOld1 <- plyr::ldply(CountyDataOld[-100])
CountyDataOld1[100,] = CountyDataOld[[100]]
# ]CountyDataOld = rbind(CountyDataOld,temp2)
temp <- strsplit(CountyDataOld1[,2], " ", fixed = TRUE)
temp <- plyr::ldply(temp)
temp[,1] = CountyDataOld1[,1]
colnames(temp) = c("County", "Individuals.Tested","Individuals.Positive", 
                   "Total.Recovered", "Total.Deaths")
CountyDataOld <- temp
CountyDataOld[-1] = sapply(CountyDataOld[-1], as.integer)
CountyDataOld <- as.data.frame(CountyDataOld) %>%
    arrange(County)%>%
    mutate(Active = Individuals.Positive - Total.Recovered - Total.Deaths,
           positivity = Individuals.Positive/Individuals.Tested,
           date = as.Date(date),
    )

temp <-c(County = "Totals",
         colSums(CountyDataOld[,2:6]),
         positivity = NA,
         date = as.character(date))
temp["positivity"] = as.integer(temp["Individuals.Positive"])/as.integer(temp["Individuals.Tested"])
CountyDataOld[101,] = temp
CountyDataOld[,2:6] = sapply(CountyDataOld[,2:6], as.integer)
CountyDataOld[,7] = sapply(CountyDataOld[,7], as.numeric)
# CountyDataOld = CountyDataOld[-100,]
write.csv(CountyDataOld, file = paste(
    'CountyData/',date,'CountyData.csv'))

#More interesting information
# tempnames =Summary[seq(7,21, 2)]
# tempnames1 = Summary[seq(26,40,2)]
# tempnames = c(tempnames, tempnames1)
# tempdata =Summary[seq(8,20, 2)]
# tempdata1 =Summary[seq(25,41, 2)]
tempnames =Summary[seq(7,17, 2)]
tempnames1 = Summary[seq(22,36,2)]
# tempnames2 = Summary[seq(40,42,2)]
tempnames2 = Summary[seq(41,43,2)]
tempnames = c(tempnames, tempnames1, tempnames2)
tempdata =Summary[seq(8,16, 2)]

tempdata1 =Summary[seq(21,35, 2)]
tempdata2 =Summary[seq(40,44, 2)]
tempdata = c(tempdata, tempdata1, tempdata2)
tempdata = plyr::ldply(as.integer(
  gsub(",","",tempdata)))
tempdata = t(tempdata)
colnames(tempdata) <- tempnames
tempdata = cbind("date"=date, tempdata)
statedata <- read.csv(paste(
  'CountyData/StateData.csv'),
                   stringsAsFactors = FALSE)
# statedata = rbind(statedata, tempdata)
if (statedata$date[nrow(statedata)] != tempdata[1]){
statedata[nrow(statedata)+1,]=tempdata
}
statedata[,-1] = sapply(statedata[,-1], as.integer)
statedata = unique(statedata)
write.csv(statedata, file = paste(
    'CountyData/StateData.csv'), 
    row.names = FALSE)
```

```{r storedata1, echo=FALSE}
##Add data to big csv files
d_county <- t(CountyDataOld)
county_names <- d_county[1,]
colnames(d_county) <- d_county[1,]
d_county = d_county[-1,]
# colnames(d_county)[1] <- "date"
d_county[1:5,] = sapply(d_county[1:5,], as.integer)
# d_county[9,] = sapply(d_county[9,], as.numeric)
# d_county <- as.data.frame(d_county)%>%
#     mutate(Total = Reduce(`+`, .))
d_county <- cbind(date=as.character(date),d_county)

Tested <- read.csv( "CountyData/CountyTests.csv", stringsAsFactors=FALSE)
Tested <- Tested[,-1]
Positive <- read.csv( "CountyData/CountyPositive.csv", stringsAsFactors=FALSE)
Positive <- Positive[,-1]
Recovered <- read.csv( "CountyData/CountyRecovered.csv", stringsAsFactors=FALSE)
Recovered <- Recovered[,-1]
Deaths <- read.csv( "CountyData/CountyDeaths.csv", stringsAsFactors=FALSE)
Deaths <- Deaths[,-1]
Active <- read.csv( "CountyData/CountyActive.csv", stringsAsFactors=FALSE)
Active <- Active[,-1]
templist = c("Tested", "Positive", "Recovered", "Deaths", "Active" )
# 
# Tested <- Tested[-nrow(Tested),]
# Positive <- Positive[-nrow(Positive),]
# Recovered <- Recovered[-nrow(Recovered),]
# Active <- Active[-nrow(Active),]
# Deaths <- Deaths[-nrow(Deaths),]


if(anydate(Tested[nrow(Tested), 1]) != date){
    Tested <- rbind(Tested, d_county["Individuals.Tested",])
    Positive <- rbind(Positive, d_county["Individuals.Positive",])
    Recovered <- rbind(Recovered, d_county["Total.Recovered",])
    Deaths <- rbind(Deaths, d_county["Total.Deaths",])
    Active <- rbind(Active, d_county["Active",])
} else{
      Tested[Tested$date == date,] = d_county["Individuals.Tested",]
    Positive[Positive$date == date,] =  d_county["Individuals.Positive",]
    Recovered[Recovered$date==date,] =  d_county["Total.Recovered",]
    Deaths[Deaths$date == date,] = d_county["Total.Deaths",]
    Active[Active$date == date,] =  d_county["Active",]
}

Tested[,-1] = sapply(Tested[,-1], as.integer)
Positive[,-1] = sapply(Positive[,-1], as.integer)
Recovered[,-1] = sapply(Recovered[,-1], as.integer)
Deaths[,-1] = sapply(Deaths[,-1], as.integer)
Active[,-1] = sapply(Active[,-1], as.integer)

trows <- nrow(Tested)
newTest <- Tested[trows,-1]-Tested[trows-1,-1]
newPos <- Positive[trows,-1]-Positive[trows-1,-1]
newRec <- Recovered[trows,-1]-Recovered[trows-1,-1]
newDeath <- Deaths[trows,-1]-Deaths[trows-1,-1]
# PerPos7 <- (Positive[trows,-1]-
#                 Positive[trows-7,-1])/
#     (Tested[trows,-1]-
#          Tested[trows-7,-1])

county_names <- colnames(Tested)
county_names <- county_names[county_names != "Pending Investigation"]
county_names <- county_names[county_names != "Pending.Investigation"]
# county_names <- county_names[-76]
county_names <- gsub(" ", ".", county_names)
county_names <- gsub("'", ".", county_names)
dates <- anydate(Tested[,1]    )
if(!("Pending.Investigation" %in% names(hospital))){
    target = which(names(hospital)== "Palo.Alto")
    hospital = cbind(hospital[,1:target,drop=F], data.frame("Pending.Investigation"=""), hospital[,(target+1):length(hospital),drop=F])
}

for(i in 2:(length(county_names))){
    i = county_names[i]
    test <- Tested[,i]
    temp = length(test)
    pos <- as.numeric(Positive[,i])
    rec <- as.numeric(Recovered[,i])
    hosp <- as.integer(hospital[,i])
    hosp = c(rep(NA, length(test)-length(hosp)),hosp)
    death <- as.numeric(Deaths[,i])
    temp <- data.frame(
        date = dates,
        Tested = test,
        Positive = pos,
        Recovered = rec,
        Hospitalized = hosp,
        Deaths = death)
    
    temp <- temp %>% 
        mutate(
            Active = Positive - Recovered - Deaths,
            New.Pos = Positive - lag(Positive),
            New.Test = Tested - lag(Tested),
            New.Rec = Recovered - lag(Recovered),
            New.Deaths = Deaths - lag(Deaths),
            Frac.Pos = New.Pos/New.Test,
            Day7.Test = Tested - lag(Tested, n=7),
            Day7.Pos = Positive - lag(Positive, n=7),
            Day7.Rec = Recovered - lag(Recovered, n=7),
            Day7.Death = Deaths - lag(Deaths, n=7),
            Day7.Active = Active - lag(Active, n=7),
            Day7.Perc = Day7.Pos/Day7.Test,
            Day14.Test = Tested - lag(Tested, n=14),
            Day14.Pos = Positive - lag(Positive, n=14),
            Day14.Rec = Recovered - lag(Recovered, n=14),
            Day14.Death = Deaths - lag(Deaths, n=14),
            Day14.Active = Active - lag(Active, n=14),
            Day14.Perc = Day14.Pos/Day14.Test,
            Not.Sick = (lag(Positive, n= 10) -
                            lag(Positive, n=28))*0.7,
            True.Active = Positive - Recovered -
                Deaths - round(Not.Sick, digits = 0)
        )
    
    assign(paste(i,"Data", sep=""), temp)
    # write.csv(temp, paste("CountyData/",county_names[i],
    # "Data.csv", sep=""))
}

write.csv(Tested, "CountyData/CountyTests.csv")
write.csv(Positive, "CountyData/CountyPositive.csv")
write.csv(Recovered, "CountyData/CountyRecovered.csv")
write.csv(Deaths, "CountyData/CountyDeaths.csv")
write.csv(Active, "CountyData/CountyActive.csv")

```
```{r filldata, results='hide'}
#####################################
#### Fill in covid19
#####################################

covid19 <- read.csv("covid19.csv", 
                    stringsAsFactors = FALSE)
## Fix an error that sometimes comes from importing from EXCEL
names(covid19)[1] <- "date"
# names(covid19)[names(covid19) == "ï..date"] <- 'date'
## Format date column correctly
covid19$date = anydate(covid19$date)
## Check to see if date already has a row in the data, if not add such a row.
if(sum(grepl(date, covid19$date)) == 0){
covid19[nrow(covid19)+1, "date"] = 
  as.character(date)
}
i=as.Date(date)
# i = as.Date("2021-11-22")
covid19[covid19$date == i,
        c("Bremer.Positive", 
          "Bremer.Recovered", "Bremer.Death")]=
    BremerData[BremerData$date==i, 
               c( "Positive", "Recovered", 
                 "Deaths")]

covid19[covid19$date == i,
        c("Butler.P", 
          "Butler.R", "Butler.D")]=
    ButlerData[ButlerData$date==i, 
               c("Positive", "Recovered", 
                 "Deaths")]
covid19[covid19$date == i,
        c("BlackHawk.P", 
          "BlackHawk.R", "BlackHawk.D")]=
    Black.HawkData[Black.HawkData$date==i, 
               c("Positive", "Recovered", 
                 "Deaths")]


    covid19[covid19$date == as.Date(date),"Total.Tested"] = 
        Tested[anydate(Tested$date) == date,"Totals"]
    
    covid19[covid19$date == as.Date(date),"positive"] = 
        Positive[anydate(Positive$date) == date,"Totals"]
    covid19[covid19$date == as.Date(date),"Recovered"] = 
        Recovered[anydate(Recovered$date) == date,"Totals"]
    covid19[covid19$date == as.Date(date),"deaths"] = Deaths[anydate(Deaths$date) == date,"Totals"]
    covid19[covid19$date == as.Date(date),"hospitalized"] = 
        statedata[anydate(statedata$date)==date, "Hospitalized"]
    covid19[covid19$date == as.Date(date),"ICU"] = 
        statedata[anydate(statedata$date)==date, "COVID.19.Patients.in.ICU"]
    covid19[covid19$date == as.Date(date),"Admit"] = statedata[anydate(statedata$date)==date, "COVID.19.Patients.Admitted.in.last.24.hrs"]

covid19[-1] = sapply(covid19[,-1], as.numeric)
covid19 = covid19 %>%
  mutate(Bremer.SS = Bremer.Positive-Bremer.Death - Bremer.Recovered,
         Butler.SS = Butler.P - Butler.R,
         BlackHawk.SS = BlackHawk.P- BlackHawk.D - BlackHawk.R)

write.csv(covid19, "covid19.csv",row.names = FALSE)

```

```{r}
qplot(date, New.Positive.Tests, data = tail(covid19, n=40))+geom_smooth(span = 0.3, se = TRUE) 

qplot(date, hospitalized, data = tail(covid19, n=80))+geom_smooth(span = 0.3, se = TRUE) 
```


```{r clean, echo = FALSE}

#######################################
##  Create Clean Dataset
#######################################

clearn <- !is.na(covid19$Recovered)
clean <- covid19[clearn,] 
clean$date = as.Date(clean$date)
cdeaths <- covid19[!is.na(covid19$deaths),]
clean <- clean %>% mutate(
  negative = Total.Tested - positive,
  New.Positive = positive - lag(positive),
  New.Negative = negative - lag(negative),
  new.hospital = hospitalized - lag(hospitalized),
  Total.Daily.Tests = Total.Tested - lag(Total.Tested),
  New.Deaths = deaths - lag(deaths),
  Still.Sick = positive - Recovered)
clean <- clean %>% mutate(
  negative = Total.Tested - positive,
  Percent.Pos = positive/Total.Tested*100,
  #np_7day = rollmean(New.Positive, k=7, align = "right", fill = NA),
  # np_7day = (positive - lag(positive, n=7))/7,
  New.Percent.Positive = New.Positive/Total.Daily.Tests,
  Percent.Hospitalized = hospitalized/Still.Sick*100
)
clean <- clean %>% mutate(
  New.Total.Tests = Total.Tests - lag(Total.Tests,n=1),
  New.Positive.Tests = Positive.Tests - lag(Positive.Tests, n=1),
  New.Bremer.Total.Tests = Bremer.Total.Tests - lag(Bremer.Total.Tests,n=1),
  New.Bremer.Positive.Tests = Bremer.Positive.Tests - lag(Bremer.Positive.Tests, n=1),
  )
```

```{r, echo = FALSE}
today <- clean[nrow(clean),]
day <- today$date
new_pos <- today$New.Positive
percent_pos <- round(today$New.Positive/today$New.Positive.Tests*100, digits = 1)

percent_pos7 <- round(clean$Week.Positivity.Rate[nrow(clean)-4]*100, digits = 1)

percent_hospital <- round(clean$Percent.Hospitalized[nrow(clean)], digits = 1)
# percent_hospital <- clean$Percent.Hospitalized[nrow(clean)]
ICU <- today$ICU
percent_ICU <- round(today$ICU/today$hospitalized*100, digits = 1)
sick <- format(today$Still.Sick, scientific = FALSE, big.mark=",")
IowaPop <- 3155070

test.positivity = round(sum(New.Positive.Tests["Total.Positive.Tests",-1])/sum(New.Tests["Total.Tests",-1])*100,digits = 1)

clean = clean %>% mutate(
  Individuals.Reported = Individuals.Tested - lag(Individuals.Tested),
  Individuals.Positive = Individuals.Postive - lag(Individuals.Postive),
  Bremer.Rep = Bremer.Individuals.Tested - lag(Bremer.Individuals.Postive))

write.csv(clean, "clean.csv")

# Adding more data to Bremer
hosp <- as.integer(hospital[,"Bremer"])
temp <- clean %>% select("date","Bremer.Positive", 
                         "Bremer.Recovered", "Bremer.Death")
names(temp) = c("date","Positive", "Recovered", "Deaths")
temp <- subset(temp, 
               !is.na(Positive) & date < "2020-07-31")%>% 
  mutate(Tested = NA, Hospitalized = NA)
temp2 <- BremerData %>% select("date","Hospitalized", "Tested","Positive", "Recovered", "Deaths")
temp = rbind(temp, temp2)
rm(temp2)
temp$Tested = as.integer(temp$Tested)
temp <- temp  %>% 
  mutate(
    Active = Positive - Recovered - Deaths,
    New.Pos = Positive - lag(Positive),
    New.Test = Tested - lag(Tested),
    New.Rec = Recovered - lag(Recovered),
    New.Deaths = Deaths - lag(Deaths),
    Frac.Pos = New.Pos/New.Test,
    # Day7.Test = Tested - lag(Tested, n=7),
    # Day7.Pos = Positive - lag(Positive, n=7),
    # Day7.Rec = Recovered - lag(Recovered, n=7),
    # Day7.Death = Deaths - lag(Deaths, n=7),
    # Day7.Active = Active - lag(Active, n=7),
    # Day7.Perc = Day7.Pos/Day7.Test,
    # Day14.Test = Tested - lag(Tested, n=14),
    # Day14.Pos = Positive - lag(Positive, n=14),
    # Day14.Rec = Recovered - lag(Recovered, n=14),
    # Day14.Death = Deaths - lag(Deaths, n=14),
    # Day14.Active = Active - lag(Active, n=14),
    # Day14.Perc = Day14.Pos/Day14.Test,
            # Not.Sick = (lag(Positive, n= 10) -
            #               lag(Positive, n=28))*0.7,
            # True.Active = Positive - Recovered -
            #   Deaths - round(Not.Sick, digits = 0)
  )

BremerData = temp
rm(temp)
write.csv(BremerData, "CountyData/BremerData.csv")

maxtoday = data.frame(colMax(clean[,-1]) == today[,-1])
temp = subset(clean, date == (as.Date(day)-365))
oldactive = temp$positive-temp$Recovered-temp$deaths
rm(temp)
week.new.pos = round(new_pos/7, digits = 1)
```

```{r}
# Vaccine Information
colTotal <- function(data) sapply(data, sum, na.rm = TRUE)
CountyLookup = as.data.frame(read.csv("CountyLookUp.csv", stringsAsFactors=F, col.names = c("RECIP_ADDRESS_COUNTY", "County")))
CompleteVaccination = read.csv("CountyData/Vaccine.csv", stringsAsFactors = F)

Vaccine = as.data.frame(read.csv("Vaccine Series Completion by Recipient County of Residence.csv", stringsAsFactors=F))
# Vaccine = sapply(Vaccine, as.integer)
# Vaccine$County <- CountyLookup[
#     match(Vaccine$RECIP_ADDRESS_COUNTY, 
#           CountyLookup$RECIP_ADDRESS_COUNTY),2]

Vaccine = Vaccine %>%
    arrange(County)

temp1 = colTotal(Vaccine[,-grep("County", colnames(Vaccine))])
Vaccine <- rbind(Vaccine, c( temp1,"Totals"))

# CompleteVaccination[,-1] = t(Vaccine$Vaccinations.Completed[-1])
CompleteVaccination[nrow(CompleteVaccination)+1,-1] = t(Vaccine$Vaccinations.Completed[-1])
CompleteVaccination[nrow(CompleteVaccination),1] = as.character(Sys.Date())

write.csv(CompleteVaccination, "CountyData/Vaccine.csv", row.names = F)


# VaccineNames = colnames(Vaccine)
# for(i in 2:ncol(Vaccine)){
#     a = VaccineNames[i]
#     temp = Vaccine[,c("County",a)]
#     temp[,a] = as.integer(temp[,a])
#     temp = rbind(c("Date",mydate), temp)
#     temp = t(temp)
#     temp1 = read.csv(paste("CountyData/",a,
#                           "Data.csv", sep=" "), header = T,
#              stringsAsFactors = FALSE)
#     rownames(temp1) = temp1[,1]
#     temp1 = temp1[,-1]
#     if(temp1[nrow(temp1),1]!= temp[2,1]){
#         temp = rbind(temp1,temp[2,])
#     }
#     assign(a,temp)
#     write.csv(temp, paste("CountyData/",a,
#      "Data.csv", sep=" "))
      # rm(temp, temp1)
#}
    
```

# Iowa COVID-19 Summary `r day`

Today we had `r format(new_pos, scientific = FALSE, big.mark=",")` new positive patients (average of `r week.new.pos` new cases per day) which is `r percent_pos`% of the newly tested people and `r test.positivity`% of all tests recorded in the last week.
`r clean$New.Deaths[nrow(clean)]` deaths in the last week. One in `r round(IowaPop/clean[nrow(clean),]$deaths, digits = 0)` people in Iowa has died of COVID-19. There are `r sick` people actively sick and `r  percent_hospital`% (`r today$hospitalized`) are hospitalized, `r percent_ICU`% (`r ICU`) of those hospitalized are in the ICU.

One year ago, we had `r format(oldactive, scientific = FALSE, big.mark=",")` actively sick and `r subset(clean, date == (day-365))$New.Positive` new positive patients per day in Iowa.

```{r}
#`r if(today$New.Deaths >=1) {"The newly recorded deaths occurred on "}` `r if(today$New.Deaths >=1) {NDeaths[NDeaths$change != 0,"NewD"]}`

no = 31
OldDeaths <- read.csv(paste0("Total Deaths",no,".csv"))
# OldDeaths <- read.csv(paste0("Total Deaths",day(mydate-1),".csv"))
OldDeaths[,1] = as.Date(OldDeaths[,1])

# weekOldDeaths <- read.csv(paste0("Total Deaths",day(mydate-7),".csv"))
# weekOldDeaths[,1] = as.Date(OldDeaths[,1])
n = "01"
NewDeaths <- read.csv(paste0("Total Deaths",n,".csv"))
# NewDeaths <- read.csv(paste0("Total Deaths",day(mydate),".csv"))
NewDeaths[,1] = as.Date(NewDeaths[,1])


NDeaths = as.data.frame(merge(NewDeaths, OldDeaths, by ="EventDateOfDeath", all=T))
NDeaths[is.na(NDeaths)] <- 0
NDeaths = NDeaths %>% mutate(change = Total.Deaths...Daily.Total.x
 -Total.Deaths...Daily.Total.y,
 NewD = paste0(EventDateOfDeath," (",change,")")) 
NDeaths[NDeaths$change != 0,"NewD"]

# NweekDeaths = as.data.frame(merge(NewDeaths, weekOldDeaths, by ="EventDateOfDeath"))
# NweekDeaths = NweekDeaths %>% mutate(change = Total.Deaths...Daily.Total.x
#                              -Total.Deaths...Daily.Total.y)
# NweekDeaths[NweekDeaths$change != 0,1]
```

#Print Graphs

```{r}
day = as.Date(day)
pactive <- qplot(date, Still.Sick, data = clean[-(1:23),], geom = "point") +
ggtitle("Known Active COVID-19 Cases in Iowa") +
  xlab("Date")+
  ylab("Active Cases")+
  geom_hline(yintercept = oldactive, color = "red") 
  # geom_text(
  #   aes(x= anydate(clean$date[1])+80, y= oldactive, 
  #       label = as.Date(day)-365, vjust = -0.5), 
  #   colour = "black")
png(paste(day,"pactive.png"))
pactive
dev.off()
 pactive
```

## Percent Positive Active Cases

```{r,echo=False}

 ppositive <-qplot(date, New.Percent.Positive, data = clean[-(1:7),], geom = c("point", "smooth")) + ylab("Fraction Positive") +
 ggtitle("Fraction of Positive COVID-19 Tests in Iowa")
 png(paste(day,"ppositive.png"))
 ppositive
 dev.off()
 ppositive

 # p7positive <-qplot(date, Week.Positivity.Rate, data = clean[-(1:20),], geom = c("point", "smooth")) + ylab("Fraction Positive") +
 # ggtitle("7 day average of Fraction of Positive COVID-19 Tests in Iowa")
 # png(paste(day,"p7positive.png"))
 # 
 # p7positive
 # dev.off()
 # p7positive
```

## Bremer County I started recording daily data on Bremer county on May 29, 2020. There will be problems with the recovered data since I have yet to get tables of data for the county and the definition of recovered changed on June 30. Thus, the following graph omits that data.

```{r, echo=FALSE}
 State_Average_Active <- today$Still.Sick/IowaPop*100000
 Bremer_Population <- county[county$COUNTY=="Bremer",][,3]
 Butler_Population <- county[county$STATE=="Iowa"&county$COUNTY=="Butler",][,3]
 BlackHawk_Population <- county[county$STATE=="Iowa"&county$COUNTY=="Black Hawk",][,3]
 pBremer <-qplot(date,
                 Bremer.SS,
                 data = tail(clean, n=180), 
#                 geom = c("point", "smooth")
                 ) +
   ggtitle("Known Active COVID-19 Cases in Bremer County") +
   geom_smooth(span = 0.3, se = TRUE)
 #    "Known Active COVID-19 Cases in Bremer County")
 #+  geom_hline(aes(yintercept=State_Average_Active)
 pButler <-qplot(date,
                 #Butler.SS/Butler_Population*100000,
                 Butler.SS,
                 data = tail(clean, n=120), geom = "point") +
   ggtitle("Known Active COVID-19 Cases in Butler County") + geom_smooth(span = 0.3, se = TRUE) 
ggtitle("Known Active COVID-19 Cases in Butler County Per 100,000 People")
#+  geom_hline(aes(yintercept=State_Average_Active)
 png(paste(day,"pButler.png"))
 pButler
 dev.off()
 pButler
 pBlackHawk <-qplot(date,
                 #Butler.SS/Butler_Population*100000,
                 BlackHawk.SS,
                 data = tail(clean, n=120), geom = "point") +
   geom_smooth(span = 0.3, se = TRUE) +
   ggtitle("Known Active COVID-19 Cases in Black Hawk County")
 png(paste(day,"pBlackHawk.png"))
 pBlackHawk
 dev.off()
 pBlackHawk

 png(paste(day,"pBremer.png"))
 pBremer
 dev.off()
 pBremer

 day7pos.Bremer <- round(
   BremerData[nrow(BremerData), ncol(BremerData)]*100, digits=1)
```
