---
title: "Iowa COVID-19 County Data"
author: "Mariah Birgen, Ph.D."
output: html_notebook
---
```{r,echo=FALSE}
require(knitr)
```

This is created to present historical County based data on COVID-19. All data was found on  https://coronavirus.iowa.gov/ . The major difference is that daily data has been cached so that it is possible to see how the numbers change as time progresses.

```{r, echo=FALSE,results="asis"}
names_list <- CountyData[,1]
names_list <- names_list[-75]
names_list <- names_list[-100]


temp1 = data.frame()
for(i in names_list){
    temp1 = data.frame()
    i1 <- gsub(" ", ".",i)
    i1 <- gsub("'", ".",i1)
test <- as.numeric(Tested[,i1])
temp = length(test)
pos <- as.numeric(Positive[,i1])
rec <- as.numeric(Recovered[,i1])
hosp <- as.integer(hospital[(nrow(hospital)-temp+1):nrow(hospital),i1])
death <- as.numeric(Deaths[,i1])
dates <- Tested[,1]    
temp <- data.frame(
    date = dates,
    Tested = test,
    Positive = pos,
    Recovered = rec,
    Hospitalized = hosp,
    Deaths = death)

temp <- temp %>% 
    mutate(
        Active = Positive - Recovered - Deaths,
        New.Pos = Positive - lag(Positive),
        New.Test = Tested - lag(Tested),
        New.Rec = Recovered - lag(Recovered),
        New.Deaths = Deaths - lag(Deaths),
        Frac.Pos = New.Pos/New.Test,
        Day7.Pos = Positive - lag(Positive, n=7),
        Day7.Rec = Recovered - lag(Recovered, n=7),
        Day7.Death = Deaths - lag(Deaths, n=7),
        Day7.Active = Active - lag(Active, n=7),
        Day7.Perc = Day7.Pos/(Tested - lag(Tested, n=7))
    )
    
    pop  <- county[
        county$STATE=="Iowa"&county$COUNTY==i,][,3]
    temp <- temp %>% 
        mutate(
            Active = Positive - Recovered - Deaths,
            PC = Active/pop*100000,
            County = i
        )
    temp1 = temp[,c("date", "County", "PC")]
    temp1$date  <- as.Date(temp1$date)
    printplot <- qplot(date, PC, color = County, data = temp1,
                       # show.legend = FALSE
                       geom = "point") +
        geom_smooth()+
      geom_hline(yintercept = 100, color = "purple")+
        ylab("Cases per 100,000") +
        labs(color = "County", title = "Active Cases Per 100,000")+
  geom_text(aes(temp1[6,"date"],100,label = "Moderate Threshold", vjust = -.5))
    
    # png(paste("CountyPlot/",i,".png"))
    # printplot
    # dev.off()
    cat('\n#', i, '\n')
    
    cat('\n')
    print(printplot)
    plot.new()
  # dev.off()
  
    cat('\n')
}
```



Note that most of the data presented uses the concept of an Active Case. Here, an active case is a patient with a positive COVID-19 test who has not recovered or passed on. The definition of recovered for most cases in Iowa became 28 days after positive test as of the end of June 2020. The author deeply appreciates that there are many people who are not feeling better 28 days after diagnosis and it might be annoying to be classified as recovered.

